---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions,echo=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE, fig.keep="all", fig.path="figure/")
```

## Loading and preprocessing the data

```{r _1_load_data}
activity <- read.csv(unz("activity.zip","activity.csv"), header=T, colClasses=c("numeric","Date","numeric"))
```

## What is mean total number of steps taken per day?

```{r _1_sum_steps}
good <- activity[!is.na(activity[[1]]),]
total_steps <- sum(good$steps)
```
The total number of steps taken: `r total_steps`

```{r _2_histogram}
#head(good)
sum_steps_per_day <- aggregate(steps ~ date, data = good, FUN = sum)
library(ggplot2)
ggplot(sum_steps_per_day, aes(x=steps)) +
  geom_histogram(binwidth=1000, colour="black", fill="white") +
  ggtitle("Total number of steps per day, ignore NA.")
```

```{r _3_total_steps_estimates}
mean_steps <- mean(sum_steps_per_day$steps)
med_steps <- median(sum_steps_per_day$steps)
```
Mean number of steps per day: `r mean_steps`
Median number of steps per day: `r med_steps`

## What is the average daily activity pattern?

```{r _1_avg_steps_plot}
mean_steps_per_interval <- aggregate(steps ~ interval, data = good, FUN = mean)
library(ggplot2)
ggplot(mean_steps_per_interval, aes(x=interval, y=steps)) + geom_line()
```

```{r _2_max_activity_interval}
max_id <- which.max(mean_steps_per_interval$steps)
max_steps_interval <- mean_steps_per_interval$interval[max_id]
```
`r max_steps_interval` interval, on average across all the days in the dataset, contains the maximum number of steps.


## Imputing missing values

```{r _1_missing_values}
total_missing <- sum(is.na(activity[[1]]))
```
The total number of missing values in the dataset: `r total_missing`.

```{r _2_fill_strategy}
fill_strategy <- function(chunk) {
  mean_steps <- mean(chunk$steps,na.rm = T);
  chunk[is.na(chunk$steps),]$steps <- mean_steps;
  chunk
}
```

```{r _3_replace_na}
filled <- do.call(rbind, lapply(split(activity,activity$interval), function(chunk) fill_strategy(chunk)))
```

```{r _4_stats}
sum_filled_steps_per_day <- aggregate(steps ~ date, data = filled, FUN = sum)
ggplot(sum_filled_steps_per_day, aes(x=steps)) +
  geom_histogram(binwidth=1000, colour="black", fill="white") +
  ggtitle("Total number of steps per day, filled.")

mean_filled_steps <- mean(sum_filled_steps_per_day$steps)
med_filled_steps <- median(sum_filled_steps_per_day$steps)
```
The mean total number of steps taken per day: `r mean_filled_steps`.
The median total number of steps taken per day: `r med_filled_steps`.

Mean estimate `r ifelse(mean_filled_steps == mean_steps, "not ", "") ` changed.
Median estimate `r ifelse(med_filled_steps == med_steps, "not ", "") ` changed.

The impact of imputing missing data on the estimates of the total daily number of steps:  
- Difference in mean number of steps per day: `r mean_filled_steps - mean_steps`.  
- Difference in median number of steps per day: `r med_filled_steps - med_steps`.

## Are there differences in activity patterns between weekdays and weekends?

```{r _1_weekend_dataset}
weekend_bool <- weekdays(filled$date) %in% c('Saturday','Sunday')
weekend_factor <- factor(as.numeric(weekend_bool), labels=c("weekday", "weekend"))
wdata <- cbind(filled, data.frame(weekend = weekend_factor))
```

```{r _2_panel_plots,echo=T}
library(ggplot2)

mean_wday_steps <- aggregate(steps ~ interval + weekend, data=wdata, FUN=mean)

p <- ggplot(mean_wday_steps, aes(x=interval, y=steps)) +
    geom_line() +
    ggtitle("Mean number of steps by 5-minute interval")

p + facet_grid(weekend ~ .)
```

